---
title: "Dual sgRNA Library CRISPRi-Seq Analysis in D39V"
author: "Julien Dénéréaz"
date: "November, 2021"
output: 
  html_document: 
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    number_sections: yes
    theme: readable
    highlight: tango
    toc_depth: 5
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache=TRUE, tidy=TRUE, fig.align = "center")
library(readxl)
library(DESeq2)
library(reshape2)
library(ggpubr)
library(visNetwork)
library(ggplot2)
library(RColorBrewer)
library(ggrepel)
library(UpSetR)
library(plotly)
library(genbankr)
library(ggtext)
library(writexl)
library(knitr)
library(dplyr)
library(patchwork)
library(tidyr)
library(parallel)
library(ggcorrplot)
library(pheatmap)
library(progress)
library(pbapply)
library(InteractiveComplexHeatmap)
library(ggtree)
library(ComplexHeatmap)
library(shiny)
library(ggbeeswarm)
library(metap)
library(stringr)
library(genbankr)
library(plyr)
library(corrplot)
library(circlize)
library(scales)
library(ggpmisc)
library(gridExtra)
library(htmlwidgets)
library(circlize)
library(gtools)
library(dplyr)
library(processx)
library(umap)

#sgRNA targets
D39V_targets <- read.table("targets_operon_manual-curation_20220401.txt", header=T)[c(1,26,27)]
colnames(D39V_targets) <- c("sgRNA", "Locus.Tag", "Gene.Name")
D39V_targets$sgRNA <- paste0("sgRNA", as.numeric(gsub("sgRNA", "", D39V_targets$sgRNA)))
D39V_targets_all <- read.table("targets_operon_manual-curation_20220401.txt", header=T)
D39V_Cogs <- read_xlsx("D39V_eggNOG_addgeneinfo.xlsx")[c(2,25)]
D39V_targets$COGs <- lapply(D39V_targets$Locus.Tag, function(operon) {
  tmp <- strsplit(operon, ",")
  cog <- c()
  for (gene in tmp) {
    g <- D39V_Cogs$COG.cat[match(gene, D39V_Cogs$locus_tag)]
    g <- unlist(strsplit(g, ""))
    if(length(g) == 0) {
      g <- NA
    }
    cog <- c(cog, g)
  }
  return(paste(cog, collapse = ","))
})
colnames(D39V_targets_all)[c(1,26,27)] <- c("sgRNA", "Locus.Tag", "Gene.Name")
#Palette
palette <- c( "#00798c","#66a182", "#edae49", "#d1495b")
mapCOG <- read.table("mapCOG.txt", sep="\t") %>%
  dplyr::rename(Abr=V1, Desc=V2)

# m1, m2: the sample means
# s1, s2: the sample standard deviations
# n1, n2: the same sizes
# m0: the null value for the difference in means to be tested for. Default is 0. 
# equal.variance: whether or not to assume equal variance. Default is FALSE. 
t.test2 <- function(m1,s1,m2,s2,n1=3,n2=3,m0=0,equal.variance=FALSE)
{
    if( equal.variance==FALSE ) 
    {
        se <- sqrt( (s1^2/n1) + (s2^2/n2) )
        # welch-satterthwaite df
        df <- ( (s1^2/n1 + s2^2/n2)^2 )/( (s1^2/n1)^2/(n1-1) + (s2^2/n2)^2/(n2-1) )
    } else
    {
        # pooled standard deviation, scaled by the sample sizes
        se <- sqrt( (1/n1 + 1/n2) * ((n1-1)*s1^2 + (n2-1)*s2^2)/(n1+n2-2) ) 
        df <- n1+n2-2
    }      
    t <- (m1-m2-m0)/se 
    dat <- c(m1-m2, se, t, 2*pt(-abs(t),df))    
    names(dat) <- c("Difference of means", "Std Error", "t", "p-value")
    return(dat[[4]]) 
}




```

# Introduction

This script analyses all of the Dual sgRNA CRISPRi libraries that was sequenced.The input is one formatted excel table, containing the read counts for each sgRNA combination in each condition.

# Dual CRISPRi-Seq with 869x869 sgRNAs

In total, 378'015 unique combinations

## Input

```{r input}
raw869 <- read.csv2("raw869_NextSeq_NovaSeq.csv", sep=",", header=T, row.names=1)



```


## DESeq2

```{r}


#Adding together Lanes from Novaseq

raw869_noLane <- melt(raw869) %>%
  separate('variable', into=c("Sequencer", "Treatment", "Replicate", "Lane"), sep="[_]")



raw869_noLane <- aggregate(.~SG1+SG2+Sequencer+Treatment+Replicate, raw869_noLane[-c(6)], sum)
raw869_noLane <- dcast(raw869_noLane, SG1+SG2~Sequencer+Treatment+Replicate)

raw869_ref <- raw869_noLane[raw869_noLane$SG1 == raw869_noLane$SG2,]
raw869_comb <- raw869_noLane[raw869_noLane$SG1 != raw869_noLane$SG2,]


raw869_colData <- data.frame(s=colnames(raw869_noLane[3:14]), row.names=colnames(raw869_noLane[3:14])) %>%
  separate(col="s", into=c("Sequencer", "Treatment", "Repl"))



row.names(raw869_comb) <- paste0(raw869_comb$SG1, "_", raw869_comb$SG2)
row.names(raw869_ref) <- raw869_ref$SG1

DDS_old <- DESeqDataSetFromMatrix(countData = raw869_comb[-c(1,2)],
                                  colData = raw869_colData,
                                  design = ~ Treatment + Sequencer) %>%
  estimateSizeFactors()
DDS_old_single <- DESeqDataSetFromMatrix(countData = raw869_ref[-c(1,2)],
                                  colData = raw869_colData,
                                  design = ~ Treatment + Sequencer) %>%
  estimateSizeFactors()



alpha <- 0.05
LFCt <- 1

DDS_old = DESeq(DDS_old)
DDS_old_single = DESeq(DDS_old_single)


DDS_old_res <- results(DDS_old, name = "Treatment_WithIPTG_vs_NoIPTG", lfcThreshold = LFCt, alpha = alpha)
DDS_old_single_res <- results(DDS_old_single, name = "Treatment_WithIPTG_vs_NoIPTG", lfcThreshold = LFCt, alpha = alpha)


df_old_res_shr <- as.data.frame(lfcShrink(DDS_old, res = DDS_old_res, type = "apeglm", coef = "Treatment_WithIPTG_vs_NoIPTG"))

df_old_single_res_shr <- as.data.frame(lfcShrink(DDS_old_single, res = DDS_old_single_res, type = "apeglm", coef = "Treatment_WithIPTG_vs_NoIPTG"))
df_old_single_res_shr$sgRNA <- row.names(df_old_single_res_shr)
df_old_res_shr$sgRNA <- row.names(df_old_res_shr)
df_old_res_shr <- separate(df_old_res_shr, 'sgRNA', into=c("SG1", "SG2"), sep="[_]", remove = T)

df_old_res_shr$SG1.targets <- D39V_targets$Gene.Name[match(df_old_res_shr$SG1, D39V_targets$sgRNA)]
df_old_res_shr$SG2.targets <- D39V_targets$Gene.Name[match(df_old_res_shr$SG2, D39V_targets$sgRNA)]


df_old_res_shr$tot <- df_old_res_shr$SG1.refLog2FC + df_old_res_shr$SG2.refLog2FC



df_old_res_shr$targets <- paste0(df_old_res_shr$SG1.targets, " | ", df_old_res_shr$SG2.targets)




df_old_res_shr$SG1.refLog2FC <- df_old_single_res_shr$log2FoldChange[match(df_old_res_shr$SG1, df_old_single_res_shr$sgRNA)]
df_old_res_shr$SG1.reflfcSE <- df_old_single_res_shr$lfcSE[match(df_old_res_shr$SG1, df_old_single_res_shr$sgRNA)]

df_old_res_shr$SG1.refpadj <- df_old_single_res_shr$padj[match(df_old_res_shr$SG1, df_old_single_res_shr$sgRNA)]
df_old_res_shr$SG2.refLog2FC <- df_old_single_res_shr$log2FoldChange[match(df_old_res_shr$SG2, df_old_single_res_shr$sgRNA)]
df_old_res_shr$SG2.reflfcSE <- df_old_single_res_shr$lfcSE[match(df_old_res_shr$SG2, df_old_single_res_shr$sgRNA)]
df_old_res_shr$SG2.refpadj <- df_old_single_res_shr$padj[match(df_old_res_shr$SG2, df_old_single_res_shr$sgRNA)]



#E(Wxy) = Wx+Wy #better for negative calling
df_old_res_shr <- transform(df_old_res_shr, expectedSum=SG1.refLog2FC+SG2.refLog2FC)







#Score Epsilon:  ε = Wxy − E(Wxy) +- lfcSE
df_old_res_shr <- df_old_res_shr %>% mutate(
  epsilonSum = log2FoldChange-expectedSum
)




#hit calling
tmp <- df_old_res_shr %>% mutate(
  interactionSum = 
    case_when(
                       (epsilonSum < -1 & baseMean > 10 & padj < 0.05 & SG1.refLog2FC > -5 & SG2.refLog2FC > -5) ~ "Negative", 
                       (epsilonSum > 1 & baseMean > 10 & padj > alpha) ~ "Positive",
                       TRUE ~ "Neutral"
                               )
)
  


df_old_res_shr$Group <- factor(df_old_res_shr$interactionSum, levels=c("Positive", "Neutral", "Negative", "Synthetic Lethal"))

df_old_res_shr$Group[which(df_old_res_shr$Group == "Negative" & df_old_res_shr$log2FoldChange < -7)] <- "Synthetic Lethal"

df_old_res_shr <- df_old_res_shr %>%
  mutate(pairs = case_when(
    SG1.refpadj < alpha & SG2.refpadj < alpha ~ "E-E",
    SG1.refpadj > alpha & SG2.refpadj > alpha ~ "NE-NE",
    SG1.refpadj < alpha | SG2.refpadj < alpha ~ "NE-E"
  )) 



```

# Mini Dual CRISPRi-Seq Analysis

## Input

```{r mini}
sgRNAList <- c(
  "sgRNA3","sgRNA348","sgRNA493","sgRNA593","sgRNA768",
  "sgRNA785","sgRNA812","sgRNA932","sgRNA187","sgRNA294",
  "sgRNA440","sgRNA500","sgRNA628","sgRNA758","sgRNA788",
  "sgRNA870","sgRNA1029","sgRNA1357","sgRNA1500"
)

rawMini <- read.csv2("NovaSeq_rc_1-2-3Mis.csv", sep=",")
rawMini <- rawMini[which(rawMini$SG1 %in% sgRNAList | rawMini$SG2 %in% sgRNAList),c(1,2,13:28)]

colnames(rawMini) <- c("SG1", "SG2", paste(rep(c("NoIPTG", "WithIPTG"), each=8, times=1), rep(c(1,2,3,4), each=2, times=2), rep(c("L1", "L2"), each=1, times=8), sep="_"))

rawMini <- separate(melt(rawMini), variable, into=c("Treatment", "Repl", "Lane"), sep="[_]")
rawMini <- aggregate(.~SG1+SG2+Treatment+Repl, rawMini[-c(5)], sum)
rawMini <- dcast(rawMini, SG1+SG2~Treatment+Repl)


```

## Formating input

We are creating a formated dataframe, with 19 \* conditions columns

```{r message=FALSE, cache=TRUE, list(width.cutoff=60), tidy=TRUE, tidy.opts=list(width.cutoff=60)}


rawMini_format <- data.frame(row.names=paste0("sgRNA", 1:1499))


for(i in sgRNAList) {
  rawMini_format[paste(i,colnames(rawMini[-c(1,2)]), sep="_")] <- 0
}

pbapply(rawMini, 1, function(row) {
  v <- c()
  if(row[[1]] %in% sgRNAList) {
    v <- paste(row[[1]], colnames(rawMini[-c(1,2)]), sep="_")
    for(colName in v) {
      rawMini_format[row[[2]], colName] <<- as.numeric(row[[gsub(paste0(row[[1]], "_"), "", colName)]])
    }
  }
  if(row[[2]] %in% sgRNAList) {
    v <- paste(row[[2]], colnames(rawMini[-c(1,2)]), sep="_")
    for(colName in v) {
      rawMini_format[row[[1]], colName] <<- as.numeric(row[[gsub(paste0(row[[2]], "_"), "", colName)]])
    }
  }
})
rawMini_format[is.na(rawMini_format)] <- 0

head(rawMini_format[1:4])

```

## Normalising

We will normalize the counts of the singles combinations (sgRNA3 with sgRNA3)

```{r message=FALSE, cache=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
lapply(sgRNAList, function(sg) {
  rawMini_format[sg, grepl(paste0(sg,"_"), colnames(rawMini_format))] <<- round(mean(as.numeric(rawMini_format[sg, !grepl(paste0(sg,"_"), colnames(rawMini_format))])))
})
rawMini_format[is.na(rawMini_format)] <- 0

#Removing the last line, which is sgRNA1500 => luciferase. The line has read counts, which is basically inserted by the two algorithms above in the apply() and lapply() function.
rawMini_format <- rawMini_format[1:1499,]

head(rawMini_format[1:4])

```

## DDS object

```{r cache=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

coldataMini <- data.frame(s=colnames(rawMini_format), row.names=colnames(rawMini_format))
coldataMini <- separate(coldataMini, col="s", into=c("Background", "Treatment", "Repl"))

DDS_mini <- DESeqDataSetFromMatrix(countData = rawMini_format, 
                                  colData = coldataMini, 
                                  design = ~ Background * Treatment + Repl)

DDS_mini <- estimateSizeFactors(DDS_mini)

DDS_mini_norm <- counts(DDS_mini, normalized = TRUE)




```

## DESeq2

```{r cache=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

alpha = 0.05
LFCt = 1
#putting reference as sgRNA1500 (luciferase)
DDS_mini$Background = relevel(DDS_mini$Background, ref="sgRNA1500")


DDS_mini = DESeq(DDS_mini)
resultsNames(DDS_mini)
DDS_mini_res <- pblapply(paste0("Background", sgRNAList[1:18]), function(sg) {
  st <- as.character(paste0(sg, ".TreatmentWithIPTG"))
  res <- results(DDS_mini, lfcThreshold = LFCt, name=st)
  res_shr <- as.data.frame(lfcShrink(DDS_mini, res = res, type = "apeglm", coef = st))
  res_shr$sgRNA=row.names(res_shr)
  return(res_shr)
})
names(DDS_mini_res) <- sgRNAList[1:18]


df_mini <- bind_rows(DDS_mini_res, .id="Background")

rawMini_refDF <- rawMini_format[grepl("sgRNA1500", colnames(rawMini_format))]

DDS_mini_ref_colData <- data.frame(s=colnames(rawMini_refDF), row.names=colnames(rawMini_refDF))
DDS_mini_ref_colData <- separate(DDS_mini_ref_colData, col="s", into=c("Background", "Treatment", "Repl"))[-c(1)]
DDS_mini_ref <- DESeqDataSetFromMatrix(countData = rawMini_refDF,
                                  colData = DDS_mini_ref_colData,
                                  design = ~ Treatment + Repl)
DDS_mini_ref <- estimateSizeFactors(DDS_mini_ref)
DDS_mini_ref <- DESeq(DDS_mini_ref)
resultsNames(DDS_mini_ref)
DDS_mini_ref_res <- results(DDS_mini_ref, lfcThreshold = LFCt, name="Treatment_WithIPTG_vs_NoIPTG", alpha=alpha)
df_mini_ref <- as.data.frame(lfcShrink(DDS_mini_ref, res = DDS_mini_ref_res, type = "apeglm", coef = "Treatment_WithIPTG_vs_NoIPTG"))
df_mini_ref$sgRNA = row.names(df_mini_ref)
df_mini_ref$targets <- D39V_targets$Gene.Name[match(df_mini_ref$sgRNA, D39V_targets$sgRNA)]

df_mini$bg.targets <- D39V_targets$Gene.Name[match(df_mini$Background, D39V_targets$sgRNA)]
df_mini$targets <- D39V_targets$Gene.Name[match(df_mini$sgRNA, D39V_targets$sgRNA)]





```


# Network for PneumoGIN

```{r}

edges <- df_old_res_shr %>%
  select(c(epsilonSum, SG1, SG2, Group, targets)) %>%
  filter(Group != "Neutral") %>%
  # filter(!grepl("tarQ|gtrB|glmU|cps2O|trmB|lysS|prs2|divIVA|pulA|rpe|miaA", targets)) %>%
  filter(epsilonSum > 1 | epsilonSum < -1) %>%
  dplyr::rename(from = SG1) %>%
  dplyr::rename(to = SG2)


nodes <- data.frame(id=unique(c(as.character(edges$from), as.character(edges$to))))
nodes$label <- D39V_targets$Gene.Name[match(nodes$id, D39V_targets$sgRNA)]
# nodes$label <- NA
nodes$locus <- D39V_targets$Locus.Tag[match(nodes$id, D39V_targets$sgRNA)]
nodes$group <- D39V_targets$COGs[match(nodes$id, D39V_targets$sgRNA)]


nodes[grepl("cps2A", nodes$label), "label"] <- "cps2A-N"
nodes[grepl("cps2F", nodes$label), "label"] <- "cps2F-N"
nodes[grepl("mltG", nodes$label), "label"] <- "greA,mpgA"
nodes[grepl("orfX", nodes$label), "label"] <- "rocS"


edges$COG <- paste0(D39V_targets$COGs[match(edges$from, D39V_targets$sgRNA)], ",", D39V_targets$COGs[match(edges$to, D39V_targets$sgRNA)])
edges$length <- round((max(abs(edges$epsilonSum))-abs(edges$epsilonSum) + 1) * 25)
# edges$color.opacity <- (abs(edges$epsilonSum)-min(abs(edges$epsilonSum)))/(max(abs(edges$epsilonSum)) - min(abs(edges$epsilonSum)))
edges$color.color <- ifelse(edges$Group=="Positive", "#00798c",
                      ifelse(edges$Group == "Negative","#edae49", "#d1495b"))

nodes %>%
  apply(2, as.character) %>%
  write.table("nodes.txt",quote = F, sep="\t", row.names=F)
write.table(edges, "edges.txt", quote=F, sep="\t", row.names=F)
```


# Exporting

```{r}
tmp <- df_old_res_shr[-which(names(df_old_res_shr) %in% c("tot", "SG1.COGs", "SG2.COGs"))]
write.csv(apply(tmp, 2, "as.character"), "20230808_res_t1.csv", quote = T)
write.table(apply(tmp, 2, "as.character"), "20231011_res_t1.txt")

```

